---
title: "Analysis for revision"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
library(dplyr)
library(tidyr)
library(tidyverse)

library(ggplot2)
library(lme4)
library(lmerTest)
library(sjPlot)
library(psych)
library(car)

```


# Import training group training data, and control group assessment trial-by-trial data for heartbeat discrimination task

```{r Import trial by trial data from training group}
df <- read.csv("data/data.csv")
data_path <- "./data/processed/"
csv_files <- list.files(path = data_path, pattern = "\\.csv$", full.names = TRUE)
data_list <- csv_files %>%
  set_names(~ tools::file_path_sans_ext(basename(.x))) %>%
  map(read_csv, show_col_types = FALSE)


# Check number of trials for each training participant
row_counts <- map_int(data_list, nrow)
print(row_counts)

```



```{r Wrangle training group}

training_df <- imap_dfr(data_list, ~ mutate(.x)) # , participant = .y

# Recode 'session' column: bl -> 1, MA -> 2, exit -> 3
training_df <- training_df %>%
  mutate(
    session = case_when(
      session == "bl"   ~ 1L,
      session == "MA"   ~ 2L,
      session == "exit" ~ 3L,
      TRUE              ~ NA_integer_
    )
  )

# Remove 'sub' and 'C' from the participant column, convert to integer, then sort
training_df <- training_df %>%
  mutate(
    Participant = str_remove(Participant, "sub"),
    Participant = str_remove(Participant, "C"),
    Participant = as.integer(Participant)
  ) %>%
  arrange(Participant)

head(training_df)


```


```{r Import trial by trial data from control group}

data_path <- "./data/assessment/control/"
csv_files <- list.files(path = data_path, pattern = "\\.csv$", full.names = TRUE)
data_list <- csv_files %>%
  set_names(~ tools::file_path_sans_ext(basename(.x))) %>%
  map(read_csv, show_col_types = FALSE)


row_counts <- map_int(data_list, nrow)
print(row_counts)

```



```{r Wrangle control group}


control_df <- imap_dfr(data_list, ~ mutate(.x)) # , participant = .y

# Recode 'session' column: bl -> 1, MA -> 2, exit -> 3
control_df <- control_df %>%
  mutate(
    session = case_when(
      session == "bl"   ~ 1L,
      session == "MA"   ~ 2L,
      session == "exit" ~ 3L,
      TRUE              ~ NA_integer_
    )
  )

# Remove 'sub' and 'C' from the participant column, convert to integer, then sort
control_df <- control_df %>%
  mutate(
    Participant = str_remove(Participant, "sub"),
    Participant = str_remove(Participant, "C"),
    Participant = as.integer(Participant)
  ) %>%
  arrange(Participant)

head(control_df)


```

```{r Import and wrangle training group assessment data}
data_path <- "./data/assessment/training/"
csv_files <- list.files(path = data_path, pattern = "\\.csv$", full.names = TRUE)
data_list <- csv_files %>%
  set_names(~ tools::file_path_sans_ext(basename(.x))) %>%
  map(read_csv, show_col_types = FALSE)
row_counts <- map_int(data_list, nrow)

print(row_counts)

training_assessment_df <- imap_dfr(data_list, ~ mutate(.x)) # , participant = .y

# Recode 'session' column: bl -> 1, MA -> 2, exit -> 3
training_assessment_df <- training_assessment_df %>%
  mutate(
    session = case_when(
      session == "bl"   ~ 1L,
      session == "MA"   ~ 2L,
      session == "exit" ~ 3L,
      TRUE              ~ NA_integer_
    )
  )

# Remove 'sub' and 'C' from the participant column, convert to integer, then sort
training_assessment_df <- training_assessment_df %>%
  mutate(
    Participant = str_remove(Participant, "sub"),
    Participant = str_remove(Participant, "C"),
    Participant = as.integer(Participant)
  ) %>%
  arrange(Participant)

head(training_assessment_df)


```


```{r Calculate summary: overall response bias, overall accuracy}

summary_training_df <- training_df %>%
  group_by(Participant) %>%
  summarise(mean_response = mean(response, na.rm = TRUE),
            mean_accuracy = mean(response==inSync, na.rm = TRUE)) %>%
  ungroup()

summary_training_assessment_df <- training_assessment_df %>%
  group_by(Participant) %>%
  summarise(mean_response = mean(response, na.rm = TRUE),
            mean_accuracy = mean(response==inSync, na.rm = TRUE)) %>%
  ungroup()


summary_control_df <- control_df %>%
  group_by(Participant) %>%
  summarise(mean_response = mean(response, na.rm = TRUE),
            mean_accuracy = mean(response==inSync, na.rm = TRUE)) %>%
  ungroup()

print("Training group - trials pooled across training sessions")
print(mean(summary_training_df$mean_response))
sum(summary_training_df$mean_response < 0.5) %>% print()
sum(summary_training_df$mean_accuracy < 0.5) %>% print()

print("Training group - trials pooled across assessment sessions")
print(mean(summary_training_assessment_df$mean_response))
sum(summary_training_assessment_df$mean_response < 0.5) %>% print()
sum(summary_training_assessment_df$mean_accuracy < 0.5) %>% print()

print("Control group - trials pooled across assessment sessions")
print(mean(summary_control_df$mean_response))
sum(summary_control_df$mean_response < 0.5) %>% print()
sum(summary_control_df$mean_accuracy < 0.5) %>% print()



```

# Table 3: Calculate chance performance

```{r Calculate d prime, AUC}

control_SDT_df <- control_df %>%
  group_by(Participant, session) %>%
  summarise(
    hits = sum(response == 1 & inSync == 1),
    misses = sum(response == 0 & inSync == 1),
    false_alarms = sum(response == 1 & inSync == 0),
    correct_rejections = sum(response == 0 & inSync == 0),

    n_target = sum(inSync == 1),
    n_nontarget = sum(inSync == 0),

    hit_rate = if_else(n_target > 0, hits / n_target, NA_real_),
    miss_rate = if_else(n_target > 0, misses / n_target, NA_real_),
    false_alarm_rate = if_else(n_nontarget > 0, false_alarms / n_nontarget, NA_real_),
    correct_rejection_rate = if_else(n_nontarget > 0, correct_rejections / n_nontarget, NA_real_),
    
    accuracy = mean(response == inSync, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    hit_rate_adj = (hits + 0.5) / (n_target + 1),
    false_alarm_rate_adj = (false_alarms + 0.5) / (n_nontarget + 1),

    z_hit = qnorm(hit_rate_adj),
    z_fa = qnorm(false_alarm_rate_adj),

    d_prime = z_hit - z_fa,
    C = -0.5 * (z_hit + z_fa),

    AUC = pnorm(d_prime / sqrt(2)),

    A_prime = case_when(
      hit_rate > false_alarm_rate ~ 0.5 + ((hit_rate - false_alarm_rate) * (1 + hit_rate - false_alarm_rate)) / (4 * hit_rate * (1 - false_alarm_rate)),
      hit_rate < false_alarm_rate ~ 0.5 - ((false_alarm_rate - hit_rate) * (1 + false_alarm_rate - hit_rate)) / (4 * false_alarm_rate * (1 - hit_rate)),
      hit_rate == false_alarm_rate ~ 0.5,
      TRUE ~ NA_real_
    )  
  ) %>%  
  arrange(Participant)

```

```{r Calculate signal detection theory metrics}

calculate_SDT_metrics <- function(df) {
  df %>%
    group_by(Participant, session) %>%
    summarise(
      hits = sum(response == 1 & inSync == 1),
      misses = sum(response == 0 & inSync == 1),
      false_alarms = sum(response == 1 & inSync == 0),
      correct_rejections = sum(response == 0 & inSync == 0),

      n_target = sum(inSync == 1),
      n_nontarget = sum(inSync == 0),

      hit_rate = if_else(n_target > 0, hits / n_target, NA_real_),
      miss_rate = if_else(n_target > 0, misses / n_target, NA_real_),
      false_alarm_rate = if_else(n_nontarget > 0, false_alarms / n_nontarget, NA_real_),
      correct_rejection_rate = if_else(n_nontarget > 0, correct_rejections / n_nontarget, NA_real_),

      accuracy = mean(response == inSync, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(
      hit_rate_adj = (hits + 0.5) / (n_target + 1),
      false_alarm_rate_adj = (false_alarms + 0.5) / (n_nontarget + 1),

      z_hit = qnorm(hit_rate_adj),
      z_fa = qnorm(false_alarm_rate_adj),

      d_prime = z_hit - z_fa,
      C = -0.5 * (z_hit + z_fa),

      AUC = pnorm(d_prime / sqrt(2)),

      A_prime = case_when(
        hit_rate > false_alarm_rate ~ 0.5 + ((hit_rate - false_alarm_rate) * (1 + hit_rate - false_alarm_rate)) / (4 * hit_rate * (1 - false_alarm_rate)),
        hit_rate < false_alarm_rate ~ 0.5 - ((false_alarm_rate - hit_rate) * (1 + false_alarm_rate - hit_rate)) / (4 * false_alarm_rate * (1 - hit_rate)),
        hit_rate == false_alarm_rate ~ 0.5,
        TRUE ~ NA_real_
      )
    ) %>%
    arrange(Participant)
}

control_SDT_df <- calculate_SDT_metrics(control_df)
training_SDT_df <- calculate_SDT_metrics(training_df)
training_assessment_SDT_df <- calculate_SDT_metrics(training_assessment_df)
```


```{r Session-wise AUC, Aprime, response accuracy, and dprime 'interoceptors'}

control_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_auc_above_0.50 = sum(AUC >= 0.50, na.rm = TRUE),
    n_auc_between_0.40_0.60 = sum(AUC >= 0.40 & AUC <= 0.60, na.rm = TRUE),
    n_auc_below_0.50 = sum(AUC < 0.50, na.rm = TRUE),
    
    n_Aprime_above_or_equal_0.5 = sum(A_prime >= 0.5, na.rm = TRUE),
    n_Aprime_below_0.5 = sum(A_prime < 0.5, na.rm = TRUE),
    n_Aprime_above_0.75 = sum(A_prime >= 0.75, na.rm = TRUE),
    n_Aprime_near_chance = sum(A_prime > .4 & A_prime < .6, na.rm = TRUE),
    
    n_accuracy_above_or_equal_0.5 = sum(accuracy >= 0.5, na.rm = TRUE),
    n_accuracy_below_0.5 = sum(accuracy < 0.5, na.rm = TRUE),

    n_dprime_above_0.75 = sum(d_prime > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime > -.25 & d_prime < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime < -0.60, na.rm = TRUE)

  )

training_assessment_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_auc_above_0.50 = sum(AUC >= 0.50, na.rm = TRUE),
    n_auc_between_0.40_0.60 = sum(AUC >= 0.40 & AUC <= 0.60, na.rm = TRUE),
    n_auc_below_0.50 = sum(AUC < 0.50, na.rm = TRUE),
    
    n_Aprime_above_or_equal_0.5 = sum(A_prime >= 0.5, na.rm = TRUE),
    n_Aprime_below_0.5 = sum(A_prime < 0.5, na.rm = TRUE),
    n_Aprime_above_0.75 = sum(A_prime >= 0.75, na.rm = TRUE),
    n_Aprime_near_chance = sum(A_prime > .4 & A_prime < .6, na.rm = TRUE),
    
    n_accuracy_above_or_equal_0.5 = sum(accuracy >= 0.5, na.rm = TRUE),
    n_accuracy_below_0.5 = sum(accuracy < 0.5, na.rm = TRUE),

    n_dprime_above_0.75 = sum(d_prime > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime > -.25 & d_prime < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime < -0.60, na.rm = TRUE)


  )

training_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_auc_above_0.50 = sum(AUC >= 0.50, na.rm = TRUE),
    n_auc_between_0.40_0.60 = sum(AUC >= 0.40 & AUC <= 0.60, na.rm = TRUE),
    n_auc_below_0.50 = sum(AUC < 0.50, na.rm = TRUE),
    
    n_Aprime_above_or_equal_0.5 = sum(A_prime >= 0.5, na.rm = TRUE),
    n_Aprime_below_0.5 = sum(A_prime < 0.5, na.rm = TRUE),
    n_Aprime_above_0.75 = sum(A_prime >= 0.75, na.rm = TRUE),
    n_Aprime_near_chance = sum(A_prime > .4 & A_prime < .6, na.rm = TRUE),
    
    n_accuracy_above_or_equal_0.5 = sum(accuracy >= 0.5, na.rm = TRUE),
    n_accuracy_below_0.5 = sum(accuracy < 0.5, na.rm = TRUE),

    n_dprime_above_0.75 = sum(d_prime > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime > -.25 & d_prime < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime < -0.60, na.rm = TRUE)

  )

```

```{r Using absolute calues of d'}
# Reviewer 2: what happens to above-chance and near-chance performance when using absolute values of d'?

control_SDT_df$d_prime_abs <- control_SDT_df$d_prime %>% abs()
training_assessment_SDT_df$d_prime_abs <- training_assessment_SDT_df$d_prime %>% abs()
training_SDT_df$d_prime_abs <- training_SDT_df$d_prime %>% abs()

control_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_dprime_above_0.75 = sum(d_prime_abs > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime_abs > -.25 & d_prime_abs < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime_abs < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime_abs < -0.60, na.rm = TRUE)

  )

training_assessment_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_dprime_above_0.75 = sum(d_prime_abs > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime_abs > -.25 & d_prime_abs < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime_abs < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime_abs < -0.60, na.rm = TRUE)
  )

training_SDT_df %>%
  group_by(session) %>%
  summarise(
    n_dprime_above_0.75 = sum(d_prime_abs > 0.75, na.rm = TRUE),
    n_dprime_near_chance = sum(d_prime_abs > -.25 & d_prime_abs < +.25, na.rm = TRUE),
    n_dprime_inverse_preference_0.75 = sum(d_prime_abs < -0.75, na.rm = TRUE),
    n_dprime_inverse_preference_0.60 = sum(d_prime_abs < -0.60, na.rm = TRUE)

  )
```

# Supplementary 1.1: Comparing heartrate during training pre- vs. post-physical activity

```{r Density plot pre- vs. post-physical activity}

# Group by Participant and condition, compute mean HR (in beats per minute)
mean_hr_df <- training_df %>%
  group_by(Participant, isExercise) %>%
  summarise(mean_hr = mean(heartRate, na.rm = TRUE), .groups = "drop")

# Calculate mean interbeat interval (in milliseconds between beats)
mean_IBI_df <- mean_hr_df %>%
  transmute(Participant,
            isExercise,
            mean_IBI_ms = 60000 / mean_hr)

# Rename isExercise column for plotting
mean_hr_df <- mean_hr_df %>%
  mutate(isExercise = factor(isExercise, 
                             levels = c(0, 1), 
                             labels = c("pre-activity", "post-activity")))
mean_IBI_df <- mean_IBI_df %>%
  mutate(isExercise = factor(isExercise, 
                             levels = c(0, 1), 
                             labels = c("pre-activity", "post-activity")))

# # Density plot heartrate (BPM)
# ggplot(mean_hr_df, aes(x = mean_hr, fill = isExercise)) +
#   geom_density(alpha = 0.6) +
#   labs(x = "Mean Heartrate", 
#        y = "Density", 
#        fill = "Physical Activity Condition") +
#   scale_fill_manual(values = c("pre-activity" = "grey70", 
#                                "post-activity" = "lightcoral")) +
#   theme_minimal()

# Density plot for interbeat interval (IBI)
ggplot(mean_IBI_df, aes(x = mean_IBI_ms, fill = isExercise)) +
  geom_density(alpha = 0.6) +
  geom_vline(xintercept = 250, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 550, linetype = "dashed", color = "black") +
  annotate("text", x = 250, y = 0.002, label = "250ms", angle = 90, vjust = -0.5) +
  annotate("text", x = 550, y = 0.002, label = "550ms", angle = 90, vjust = -0.5) +
  labs(x = "Mean Interbeat Interval (ms)", 
       y = "Density", 
       fill = "Physical Activity Condition") +
  scale_fill_manual(values = c("pre-activity" = "grey70", 
                               "post-activity" = "lightcoral")) +
  coord_cartesian(xlim = c(0, 1100)) +
  theme_minimal()

# Save figure
ggsave(
  filename = "density_mean_IBI.png",
  width = 7, height = 5, units = "in",
  dpi = 300
)

```


Heart rate does not significantly differ during the pre- vs. post- physical activity trials
```{r t-test comparing HR in pre- vs. post-physical activity}

# Wide format: one row per Participant, columns for pre and post
wide_hr <- mean_hr_df %>%
  tidyr::pivot_wider(names_from = isExercise,
                     values_from = mean_hr,
                     names_prefix = "cond_")

# Paired t-test
t.test(mean_hr ~ isExercise, data = mean_hr_df, paired = TRUE)

```

```{r LME pre- vs. post-activity HR, with random effects}

temp_df <- training_df %>%
  mutate(
    Participant = factor(Participant),
    day = factor(day),
    isExercise = factor(isExercise, levels = c(0, 1), labels = c("pre", "post"))
  )

temp_df <- temp_df %>% merge(df %>% filter(Condition=='active') %>% dplyr::select(Participant, Age, Sex))


# Random intercepts for Participant
lmm1 <- lmer(heartRate ~ isExercise + Age + Sex + (1 | Participant), data = temp_df)

# Include random intercepts for Participant and day
lmm2 <- lmer(heartRate ~ isExercise + Age + Sex + (1 | Participant) + (1 | day), data = temp_df)

# Include random slopes for exercise within participants:
lmm3 <- lmer(heartRate ~ isExercise + Age + Sex + (1 + isExercise | Participant), data = temp_df)

lmm4 <- lmer(heartRate ~ isExercise + Age + Sex + (1 + isExercise | Participant) + (1 | day), data = temp_df)

anova(lmm1, lmm2)

anova(lmm1, lmm3)

anova(lmm3, lmm4)

summary(lmm4)



# Careful, this takes a long time to run!
# tab_model(lmm4, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

```

# Test-retest reliability for conventional task measures

## Hearbeat discrimination 
```{r}

# Pivot to wide format, remove missing values
accuracy_wide <- control_SDT_df %>%
  dplyr::select(Participant, session, accuracy) %>%
  pivot_wider(names_from = session, values_from = accuracy, names_prefix = "session_")
dprime_wide <- control_SDT_df %>%
  dplyr::select(Participant, session, d_prime) %>%
  pivot_wider(names_from = session, values_from = d_prime, names_prefix = "session_")

accuracy_wide_clean <- accuracy_wide %>% drop_na()
dprime_wide_clean <- dprime_wide %>% drop_na()

# ICC for accuracy
icc_accuracy <- ICC(accuracy_wide_clean %>% dplyr::select(starts_with("session_")))
print(icc_accuracy)

# ICC for d_prime
icc_dprime <- ICC(dprime_wide_clean %>% dplyr::select(starts_with("session_")))
print(icc_dprime)

```

## Heartbeat counting 

```{r HCT ICC control}
control_HCT_df <- read.csv("data/data.csv") %>%
  filter(Condition == "control") %>%
  dplyr::select(HBT_BL, HBT_Mid, HBT_FL)

# Calculate ICC using psych::ICC
icc_results <- ICC(control_HCT_df)

print(icc_results)

```

```{r HCT ICC training}

training_HCT_df <- read.csv("data/data.csv") %>%
  filter(Condition == "active") %>%
  dplyr::select(HBT_BL, HBT_Mid, HBT_FL)

# Calculate ICC
icc_results <- ICC(training_HCT_df)
print(icc_results)

```

# Supplementary 8.3: does anxiety reduction simply relate to getting more positive feedback from correct trials?

```{r Import data, calculate mean correct trials}
df <- read.csv("data/data.csv")

# Harmonise participant identifiers
df$Participant <- ifelse(df$Condition == "control", paste0(df$Participant, "C"), df$Participant)


accuracy_df <- training_df %>%
  mutate(isCorrect = if_else(inSync == response, 1, 0)) %>%
  group_by(Participant) %>%
  summarise(mean_isCorrect = mean(isCorrect), .groups = "drop") %>%
  mutate(Participant = as.character(Participant))


temp_df <- accuracy_df %>%
  left_join(
    df %>%
      dplyr::select(Participant, State_BL, Trait_BL, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, Age, Sex),  # 👈 specify columns here
    by = c("Participant")
  )

```

```{r LME to predict change in state anxiety}
model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + State_BL + mean_isCorrect + Age + Sex + (1 | Sex), 
               data    = temp_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```

```{r LME to predict change in trait anxiety}
model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + Trait_BL + mean_isCorrect + Age + Sex + (1 | Sex), 
               data    = temp_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```


