---
title: "Interoceptive training - analysis script"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r Import packages}
library(dplyr)
library(tidyverse)
library(optLog) # Requires github installation (see repo kforthman/optLog)
library(car)
library(lme4)
library(lmerTest)
library(emmeans)
library(jtools)
library(sjPlot)
library(mixedpower)

library(ridge)
library(mediation)

library(ggplot2)
library(GGally)
library(effects)
library(gridExtra)
library(lattice)

```

# Load in data, wrangle dataframes
```{r Load csv files}
df <- read.csv("data/data.csv")

# Concatenated data parameter estimates
params_df <- read_csv("results/model_fits/IP1_IP2_pS/combine_fits.csv")
params_df$Group <- "training"
params_df$Var1 <- gsub("\\.out", "", params_df$Var1)

control_all <- read.csv("results/model_fits_assessment/IP1_IP2_pS/control/all/combine_fits.csv")
control_all$Group <- "control"
control_all$Var1 <- gsub("\\.out", "", control_all$Var1)

# Derived delta IP2 estimates
end_precision_training <- read_csv("results/end_precision_TRAINING.csv")
end_precision_training$IP2_change <- end_precision_training$IP2_end - end_precision_training$IP2
end_precision_training$Group <- "training"

end_precision_control <- read_csv("results/end_precision_CONTROL.csv")
end_precision_control$IP2_change <- end_precision_control$IP2_end - end_precision_control$IP2
end_precision_control$Group <- "control"

# Assessment session parameter estimates
training_B <- read.csv("results/model_fits_assessment/IP1_IP2_pS/training/bl/combine_fits.csv")
training_M <- read.csv("results/model_fits_assessment/IP1_IP2_pS/training/MA/combine_fits.csv")
training_F <- read.csv("results/model_fits_assessment/IP1_IP2_pS/training/exit/combine_fits.csv")

control_B <- read.csv("results/model_fits_assessment/IP1_IP2_pS/control/bl/combine_fits.csv")
control_M <- read.csv("results/model_fits_assessment/IP1_IP2_pS/control/MA/combine_fits.csv")
control_F <- read.csv("results/model_fits_assessment/IP1_IP2_pS/control/exit/combine_fits.csv")

```

```{r MAIA Trusting subscale: rectify raw scores which were incorrectly coded}
# Divide all scores above maximum by 3 (number of items in Trusting sub-scale)
df$trustingFL[df$trustingFL > 5.00] <- df$trustingFL[df$trustingFL > 5.00] / 3

# Re-calculate the change score
df$Change_MAIA_Trusting_FLmBL <- df$trustingFL - df$trustingBL

# Save useful column names
MAIA_BL_cols <- c('NoticingBL', 'not_distractingBL', 'not_worryingBL', 'attention_regulationBL','emotional_awarenessBL', 'self_regulationBL', 'body_listeningBL', 'trustingBL')
MAIA_FL_cols <- c('NoticingFL','not_distractingFL','not_worryingFL','attention_regulationFL','emotional_awarenessFL','self_regulationFL','body_listeningFL','trustingFL')

# Re-calculate the average BL, FL, and change scores
df$MAIA_average_BL <- df[MAIA_BL_cols] %>% rowMeans()
df$MAIA_average_F <- df[MAIA_FL_cols] %>% rowMeans()
df$MAIA_average_change <- df$MAIA_average_F - df$MAIA_average_BL
```

```{r Harmonise subject identifiers}

# Append 'C' to participant numbers in control group
df$Participant <- ifelse(df$Condition == "control", paste0(df$Participant, "C"), df$Participant)

# Training group
update_participant_column <- function(df) {
  
  # Rename participant column if needed
  colnames(df)[colnames(df) == "Var1"] <- "Participant"
  
  df$Participant <- sub("^sub", "", df$Participant) %>%
                      str_replace("^0+", "")
  df <- df[, !(names(df) %in% c("F"))]
  return(df)
}

params_df <- update_participant_column(params_df)
training_B <- update_participant_column(training_B)
training_M <- update_participant_column(training_M)
training_F <- update_participant_column(training_F)
end_precision_training <- update_participant_column(end_precision_training)

# Control group
update_participant_column <- function(df) {
  
  # Rename participant column if needed
  colnames(df)[colnames(df) == "Var1"] <- "Participant"
  
  df$Participant <- sub("^sub", "", df$Participant) %>%
                      str_replace("^0+", "") %>%
                      str_c("C")
  df <- df[, !(names(df) %in% c("F"))]
  return(df)
}

control_all <- update_participant_column(control_all)
control_B <- update_participant_column(control_B)
control_M <- update_participant_column(control_M)
control_F <- update_participant_column(control_F)
end_precision_control <- update_participant_column(end_precision_control)
  
```

```{r Make a training + control combined dataframe}

# Merge delta IP2 values
params_df <- merge(params_df, end_precision_training[, c("Participant", "IP2_change")], by = "Participant", all.x = TRUE)
control_all <- merge(control_all, end_precision_control[, c("Participant", "IP2_change")], by = "Participant", all.x = TRUE)

combined_df <- rbind(params_df, control_all)

```

```{r Helpful column names for later}
baseline_cols <- c("Trait_BL", "State_BL")
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
```


```{r Define function to log-transform non-normal variables - used in subsequent analysis}


param_cols <- c("IP1", "IP2", "pS")

# Can be changed to power transform if preferred
applyOptLogTransform <- function(dataframe, columns) {
  transformed_data <- optLogTransform(data.matrix(dataframe[, columns]), type = "log", scaled = FALSE, skew_thresh = .5)
  
  for (i in seq_along(columns)) {
    dataframe[, columns[i]] <- transformed_data$data[, i]
    hist(transformed_data$data[, i])
  }
  
  return(dataframe)
}

```


```{r Import heart rate measures}


import_HR <- function(directory_path) {
  # Get a list of all CSV files in the directory
  csv_files <- list.files(directory_path, pattern = "\\.csv$", full.names = TRUE)
  # Initialise empty heartrate dataframe
  HR_df <- data.frame(Participant = character(), mean_heartRate = numeric(), mean_heartRateSD = numeric(), stringsAsFactors = FALSE)

  # Loop through each CSV file
  for (file in csv_files) {
    temp_df <- read.csv(file)
    # Calculate the mean of the heartRate column
    mean_hr <- mean(temp_df$heartRate, na.rm = TRUE)
    # Calculate the mean of the heartRateSD column
    mean_hr_sd <- mean(temp_df$heartRateSD, na.rm = TRUE)
    # Append new row to HR_df
    participant_value <- temp_df$Participant[1]
    new_row <- data.frame(Participant = participant_value, mean_heartRate = mean_hr, mean_heartRateSD = mean_hr_sd, stringsAsFactors = FALSE)
    HR_df <- bind_rows(HR_df, new_row)
  }
  return(HR_df)
}

# Import heartrate data, harmonise subject IDs
training_HR <- import_HR("data/processed")
training_HR$Participant <- sub("sub", "", training_HR$Participant)
training_HR$Participant <- gsub("^0", "", training_HR$Participant)

control_HR <- import_HR("data/assessment/control")
control_HR$Participant <- sub("sub", "", control_HR$Participant)
control_HR$Participant <- sub("sub", "", control_HR$Participant)
control_HR$Participant <- gsub("^0", "", control_HR$Participant)

combined_HR <- rbind(training_HR,control_HR)

combined_HR
```

```{r Import HR and HRV averaged within each assessment session}


import_HR_session <- function(directory_path) {
  csv_files <- list.files(directory_path, pattern = "\\.csv$", full.names = TRUE)
  # Initialise dataframe
  HR_session_df <- data.frame(Participant = character(), session = character(),
                              mean_heartRate = numeric(), mean_heartRateSD = numeric(),
                              stringsAsFactors = FALSE)
  for (file in csv_files) {
    temp_df <- read.csv(file)
    # Check columns exist
    if (!all(c("Participant", "session", "heartRate", "heartRateSD") %in% names(temp_df))) {
      warning(paste("Missing required columns in file:", file))
      next
    }
    # Group by Participant and session, compute means
    summary_df <- temp_df %>%
      group_by(Participant, session) %>%
      summarise(mean_heartRate = mean(heartRate, na.rm = TRUE),
                mean_heartRateSD = mean(heartRateSD, na.rm = TRUE),
                .groups = "drop")
    HR_session_df <- bind_rows(HR_session_df, summary_df)
  }

  # Recode session from label to numeric
  HR_session_df <- HR_session_df %>%
    mutate(session = case_when(
      session == "bl"   ~ 1,
      session == "MA"   ~ 2,
      session == "exit" ~ 3,
      TRUE              ~ NA_real_
    ))

  return(HR_session_df)
}

# Import datasets
training_HR_session <- import_HR_session("data/assessment/training")
control_HR_session  <- import_HR_session("data/assessment/control")

# Harmonise participant IDs
clean_participant_id <- function(id_vec) {
  id_vec <- sub("sub", "", id_vec)
  id_vec <- gsub("^0+", "", id_vec)
  return(id_vec)
}

training_HR_session$Participant <- clean_participant_id(training_HR_session$Participant)
control_HR_session$Participant  <- clean_participant_id(control_HR_session$Participant)

# Combine datasets, sort by participant then session
combined_HR_session <- bind_rows(training_HR_session, control_HR_session) %>%
  arrange(as.numeric(Participant), session)

combined_HR_session <- combined_HR_session %>%
  mutate(session = case_when(
    session == 1 ~ "baseline",
    session == 2 ~ "midpoint",
    session == 3 ~ "final",
    TRUE         ~ NA_character_
  ))

combined_HR_session <- combined_HR_session %>%
  rename(Time = session)

combined_HR_session
```

```{r Import HR and HRV averaged over all assessment trials}



import_HR_assessment_all <- function(directory_path) {
  # Get a list of all CSV files in the directory
  csv_files <- list.files(directory_path, pattern = "\\.csv$", full.names = TRUE)
  HR_assessments_df <- data.frame(Participant = character(),
                              mean_heartRate = numeric(), mean_heartRateSD = numeric(),
                              stringsAsFactors = FALSE)

  for (file in csv_files) {
    temp_df <- read.csv(file)
    
    # Check columns exist
    if (!all(c("Participant", "session", "heartRate", "heartRateSD") %in% names(temp_df))) {
      warning(paste("Missing required columns in file:", file))
      next
    }

    # Group by Participant, compute means
    summary_df <- temp_df %>%
      group_by(Participant) %>%
      summarise(mean_heartRate = mean(heartRate, na.rm = TRUE),
                mean_heartRateSD = mean(heartRateSD, na.rm = TRUE),
                .groups = "drop")
    HR_assessments_df <- bind_rows(HR_assessments_df, summary_df)
  }
  return(HR_assessments_df)
}

# Import datasets
training_HR_assessment <- import_HR_assessment_all("data/assessment/training")
control_HR_assessment  <- import_HR_assessment_all("data/assessment/control")

# Harmonise participant IDs
clean_participant_id <- function(id_vec) {
  id_vec <- sub("sub", "", id_vec)
  id_vec <- gsub("^0+", "", id_vec)
  return(id_vec)
}

training_HR_assessment$Participant <- clean_participant_id(training_HR_assessment$Participant)
control_HR_assessment$Participant  <- clean_participant_id(control_HR_assessment$Participant)

# Combine training and control groups
combined_HR_assessment <- bind_rows(training_HR_assessment, control_HR_assessment) %>%
  arrange(as.numeric(Participant))

combined_HR_assessment
```

```{r Import heartbeat counting assessment HR data}

HCT_HR_session <- read.csv("data/assessment/HCT_HR_session.csv")

# Harmonise coding and column names
HCT_HR_session <- HCT_HR_session %>%
  mutate(session = case_when(
    session == 1 ~ "baseline",
    session == 2 ~ "midpoint",
    session == 3 ~ "final",
    TRUE         ~ NA_character_
  ))
names(HCT_HR_session)[names(HCT_HR_session) == 'session'] <- 'Time'

```

# Analysis

# Interoceptive Accuracy

## Change in conventional interoceptive task measures across groups

```{r Change in heartbeat tracking task accuracy}


temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', 'HBT_BL', 'HBT_Mid', 'HBT_FL')]

# Into long form
long_df <- pivot_longer(temp_df, cols = starts_with("HBT"), names_to = "Time", values_to = "HBT")

# Rename timepoint names 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "HBT_BL" ~ "baseline",
    Time == "HBT_Mid" ~ "midpoint",
    Time == "HBT_FL" ~ "final",
    TRUE ~ Time
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge with HCT assessment HR and HRV data
long_df <- long_df %>%
  left_join(HCT_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "midpoint", "final"))

# Fit model
model <- lmer(formula = HBT ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")

# Change over time in each group
emm1 = emmeans(model, specs = consec ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")

# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="holm")

contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0, 0, 0),
  "Training-Control @ midpoint" = c(0, 0, -1, 1, 0, 0),
  "Training-Control @ final"    = c(0, 0, 0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))
eff_size(emm2, sigma = sigma(model), edf = df.residual(model))
emm3 = emmeans(model, specs = ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
eff_size(emm3, sigma = sigma(model), edf = df.residual(model))


```
```{r Power analysis using simulation methods}

SESOI <- fixef(model) * .75 # specify SESOI, 75% of effect sizes found

power_df <- long_df %>%
  mutate(subject = as.numeric(factor(Participant)))

power_SESOI <- mixedpower(model = model, data = power_df,
                              fixed_effects = c("Group", "Time"),
                              simvar = "subject", steps = c(54),
                              critical_value = 2, n_sim = 500,
                              SESOI = SESOI, databased = T, maxCores = 10)

power_SESOI

```


```{r Change in heartbeat discrimination task d’ }

temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', 'Dprime_BL', 'Dprime_ML', 'Dprime_F')]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("Dprime"), names_to = "Time", values_to = "Dprime")

# Rename timepoint names 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "Dprime_BL" ~ "baseline",
    Time == "Dprime_ML" ~ "midpoint",
    Time == "Dprime_F" ~ "final",
    TRUE ~ Time
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "midpoint", "final"))

long_df

model <- lmer(formula = Dprime ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")

# Change over time in each group
emm1 = emmeans(model, specs = consec ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")


# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="holm")
contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0, 0, 0),
  "Training-Control @ midpoint" = c(0, 0, -1, 1, 0, 0),
  "Training-Control @ final"    = c(0, 0, 0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)

print("----------------------- Normality of residuals -----------------------")

# residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))
eff_size(emm2, sigma = sigma(model), edf = df.residual(model))
emm3 = emmeans(model, specs = ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
eff_size(emm3, sigma = sigma(model), edf = df.residual(model))

```

```{r Power analysis using simulation methods}

SESOI <- fixef(model) * .75 # specify SESOI, 75% of effect sizes found


power_df <- long_df %>%
  mutate(subject = as.numeric(factor(Participant)))

power_SESOI <- mixedpower(model = model, data = power_df,
                              fixed_effects = c("Group", "Time"),
                              simvar = "subject", steps = c(54),
                              critical_value = 2, n_sim = 500,
                              SESOI = SESOI, databased = T, maxCores = 10)

power_SESOI

```

```{r Change in heartbeat discrimination task C (Supplementary 2.3)}

temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', 'Cvalue_BL', 'Cvalue_ML', 'Cvalue_F')]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("Cvalue"), names_to = "Time", values_to = "Cvalue")

# Rename timepoint names 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "Cvalue_BL" ~ "baseline",
    Time == "Cvalue_ML" ~ "midpoint",
    Time == "Cvalue_F" ~ "final",
    TRUE ~ Time  
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "midpoint", "final"))

long_df

model <- lmer(formula = Cvalue ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 + Group | Participant), 
               data    = long_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- EMMEANS -----------------------")

# Change over time in each group
emm1 = emmeans(model, specs = consec ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="Holm")

# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="none")

contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0, 0, 0),
  "Training-Control @ midpoint" = c(0, 0, -1, 1, 0, 0),
  "Training-Control @ final"    = c(0, 0, 0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)


print("----------------------- Normality of residuals -----------------------")

# residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))

```


```{r D-prime effects do not change when using absolute values of d-prime - Supplementary 2.6}


temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', 'Dprime_BL', 'Dprime_ML', 'Dprime_F')]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("Dprime"), names_to = "Time", values_to = "Dprime")

# Rename timepoint names 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "Dprime_BL" ~ "baseline",
    Time == "Dprime_ML" ~ "midpoint",
    Time == "Dprime_F" ~ "final",
    TRUE ~ Time  # Keep other values unchanged
  ))

# Rename columns
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "midpoint", "final"))

# For reviewer 2: use absolute values of d'
long_df$Dprime <- abs(long_df$Dprime)

long_df



model <- lmer(formula = Dprime ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


# Change over time in each group
emm1 = emmeans(model, specs = consec ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")

# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="none")

contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0, 0, 0),
  "Training-Control @ midpoint" = c(0, 0, -1, 1, 0, 0),
  "Training-Control @ final"    = c(0, 0, 0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))


```

# Anxiety
```{r Baseline anxiety normal distribution tests}

shapiro.test(df$Trait_BL)
shapiro.test(df$State_BL)

```

```{r Change in state anxiety}

temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', "State_BL", "State_FL")]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("State"), names_to = "Time", values_to = "State")

# Rename timepoints 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "State_BL" ~ "baseline",
    Time == "State_ML" ~ "midpoint",
    Time == "State_FL" ~ "final",
    TRUE ~ Time
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "final"))

long_df


model <- lmer(formula = State ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)
summary(model)

# vif(model)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


# Change over time in each group
emm1 = emmeans(model, specs = consec ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")

# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="none")

contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0),
  "Training-Control @ final" = c(0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)



print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))
eff_size(emm2, sigma = sigma(model), edf = df.residual(model))

```

```{r Change in state anxiety with response variable log-transformed}

temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', "State_BL", "State_FL")]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("State"), names_to = "Time", values_to = "State")

# Rename timepoints 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "State_BL" ~ "baseline",
    Time == "State_ML" ~ "midpoint",
    Time == "State_FL" ~ "final",
    TRUE ~ Time
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "final"))

# Transform the response variable does make residuals normally distributed
long_df <- applyOptLogTransform(long_df, c("State"))

long_df



model <- lmer(formula = State ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)

summary(model)

# vif(model)




tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


# Change over time in each group
emm1 = emmeans(model, specs = pairwise ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="none")

# Group differences at each timepoint
# emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
# summary(emm2, adjust="none")
# 
# contrast_list <- list(
#   "Training-Control @ baseline" = c(-1, 1, 0, 0),
#   "Training-Control @ final" = c(0, 0, -1, 1)
# )
# group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
# summary(group_contrasts)


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))

```


```{r Change in trait anxiety}

temp_df <- df[,c('Condition', 'Participant', 'Age', 'Sex', "Trait_BL", "Trait_FL")]

# Convert to long form
long_df <- pivot_longer(temp_df, cols = starts_with("Trait"), names_to = "Time", values_to = "Trait")

# Rename timepoints 
long_df <- long_df %>%
  mutate(Time = case_when(
    Time == "Trait_BL" ~ "baseline",
    Time == "Trait_ML" ~ "midpoint",
    Time == "Trait_FL" ~ "final",
    TRUE ~ Time
  ))

# Rename 'Condition' column to 'Group'
names(long_df)[names(long_df) == "Condition"] <- "Group"

# Recode group: control = -1, training = 1
long_df$Group <- ifelse(long_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

# Reorder timepoints for emmgrid
long_df$Time <- factor(long_df$Time, levels = c("baseline", "final"))

model <- lmer(formula = Trait ~ Group * Time + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Participant), 
               data    = long_df)
summary(model)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


# Change over time in each group
emm1 = emmeans(model, specs = pairwise ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")

# Group differences at each timepoint
emm2 = emmeans(model, specs =  pairwise ~  Group*Time, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm2, adjust="none")

contrast_list <- list(
  "Training-Control @ baseline" = c(-1, 1, 0, 0),
  "Training-Control @ final" = c(0, 0, -1, 1)
)
group_contrasts <- contrast(emm2, contrast_list, adjust = "holm")
summary(group_contrasts)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))
eff_size(emm2, sigma = sigma(model), edf = df.residual(model))


```

# Computational modelling

```{r Parameter estimates normal distribution tests}

# In training group only
shapiro.test(params_df$IP1)
shapiro.test(params_df$IP2)
shapiro.test(params_df$pS)
shapiro.test(end_precision_training$IP2_change)

# In control group only
shapiro.test(control_all$IP1)
shapiro.test(control_all$IP2)
shapiro.test(control_all$pS)
shapiro.test(end_precision_control$IP2_change)

# Both groups pooled
shapiro.test(combined_df$IP1)
shapiro.test(combined_df$IP2)
shapiro.test(combined_df$pS)
shapiro.test(combined_df$IP2_change)

```

```{r Figure 3}

temp_df <- combined_df[combined_df$Group == "training", c("IP1", "IP2", "pS", "IP2_change")]

GGally::ggpairs(data = temp_df, 
                upper = list(continuous = GGally::wrap(ggally_cor, stars = F)),
                columnLabels = c("IP1","IP2","pS","delta_IP2"))

```

```{r Figure 4}

temp_df <- combined_df[combined_df$Group == "control", c("IP1", "IP2", "pS", "IP2_change")]

GGally::ggpairs(data = temp_df, 
                upper = list(continuous = GGally::wrap(ggally_cor, stars = F)),
                columnLabels = c("IP1","IP2","pS","delta_IP2"))

```

## Relationship with heartrate, heartrate variability and conventional interoceptive task measures 

```{r Prepare data with heartrate measures}
temp_df <- merge(combined_df, combined_HR, by = "Participant", all.x = TRUE)
temp_df <- merge(temp_df, df[, c('Participant','Age', 'Sex')], by = "Participant", all.x = TRUE)

# temp_df <- temp_df %>% filter(Group == "training") 

temp_df

shapiro.test(temp_df$mean_heartRate)
shapiro.test(temp_df$mean_heartRateSD)
shapiro.test(temp_df$IP1)
shapiro.test(temp_df$IP2)
shapiro.test(temp_df$pS)


# temp_df[c("Participant","mean_heartRate")]

cor.test(temp_df$IP1, temp_df$mean_heartRate, method = "pearson")
cor.test(temp_df$IP1, temp_df$mean_heartRate, method = "spearman")

plot(temp_df$IP1, temp_df$mean_heartRate)

```

```{r Multiple regression to predict IP1 in both groups pooled (supplementary 2.1)}

model <- lm(formula = IP1 ~ mean_heartRate + mean_heartRateSD + IP2 + pS + Age + Sex, 
               data    = temp_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE, digits = 3)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r Effects plot IP1 and mean heartrate (Figure 7)}

trellis.par.set(
  fontsize = list(
    text = 18,  
    points = 12 
  )
)

est1 <- Effect("mean_heartRate", model, residuals=TRUE, KR=TRUE)
plt1 <- plot(est1,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,      
             partial.residuals = list(
               smooth = TRUE,
               pch = 16,
               col = "black",
               cex = 0.7,
               lty = "dashed"
             ),
             axes = list(
               x = list(mean_heartRate = list(lab = "Mean heart rate (BPM)")),
               y = list(lab = bquote("Afferent signal precision (" ~ IP[1] ~ ")")),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)

# Save figure
figures_path <- "figures_revision"

# Create folder if it doesn't exist
if (!dir.exists(figures_path)) {
  dir.create(figures_path, recursive = TRUE)
}
png(file.path(figures_path, "Figure_7.png"), width = 600, height = 400)
grid.arrange(plt1, ncol = 1)
dev.off()


```


```{r Multiple regression to predict IP1 in training group only (supplementary 2.2)}

temp_df <- temp_df %>% filter(Group == "training")

model <- lm(formula = IP1 ~ mean_heartRate + mean_heartRateSD + IP2 + pS + Age + Sex, 
               data    = temp_df)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE, digits = 3)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)


```


```{r Correlations between parameters and change in signal detection d' (training group only)}

temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE) %>% 
  filter(Group == "training")
# temp_df

# delta IP2 and change in d'
cor.test(temp_df$IP2_change, (temp_df$Dprime_F - temp_df$Dprime_BL), method = "spearman")

# IP1 and change in d'
cor.test(temp_df$IP1, (temp_df$Dprime_F - temp_df$Dprime_BL), method = "pearson")

# IP1 and delta IP2
cor.test(temp_df$IP2_change, temp_df$IP1, method = "spearman")

```

```{r Correlations between parameters and change in signal detection d' (both groups pooled)}

temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)
# temp_df

# delta IP2 and change in d'
cor.test(temp_df$IP2_change, (temp_df$Dprime_F - temp_df$Dprime_BL), method = "spearman")

# IP1 and change in d'
cor.test(temp_df$IP1, (temp_df$Dprime_F - temp_df$Dprime_BL), method = "pearson")

# IP1 and delta IP2
cor.test(temp_df$IP2_change, temp_df$IP1, method = "spearman")

```

```{r Mediation analysis (IP1 -> delta_IP2 -> change in d-prime accounting for baseline d-prime)}
# Prepare data for mediation
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)

mediation_df <- temp_df[c(param_cols, "IP2_change", "Dprime_BL", "Dprime_F")] %>% na.omit()

# Model for mediator (baseline mediator ~ X)
mod_m <- lm(IP2_change ~ IP1, data = mediation_df)  

# Model for outcome (Dprime-final controlling for Dprime-baseline and mediator)
mod_y <- lm(Dprime_F ~ IP1 + IP2_change + Dprime_BL, data = mediation_df)

# Mediation analysis
set.seed(67)
med <- mediation::mediate(model.m = mod_m,
                          model.y = mod_y,
                          treat   = "IP1",
                          mediator= "IP2_change",
                          sims    = 2000,
                          boot    = TRUE,
                          boot.ci.type = "bca")

summary(med)

```

## Computational parameter estimates from separate assessment sessions
Convert assessment data to long format, recode group if desired:
```{r}
# Add 'time' column to each dataframe, categorical coding
active_assessments <- Map(function(df, time) {
  df$time <- time
  return(df)
}, list(training_B, training_M, training_F), c("baseline","midpoint","final"))

# Combine the dataframes vertically
long_active_assessments <- bind_rows(active_assessments)

# Sort the combined dataframe by 'participant'
long_active_assessments <- long_active_assessments %>% arrange(Participant)

# Add 'group' column with value 'training' for all rows
long_active_assessments$group <- 'training'

# Control assessments:
# Add 'time' column to each dataframe
control_assessments <- Map(function(df, time) {
  df$time <- time
  return(df)
}, list(control_B, control_M, control_F), c("baseline","midpoint","final"))

# Combine the dataframes vertically
long_control_assessments <- bind_rows(control_assessments)

# Sort the combined dataframe by 'participant'
long_control_assessments <- long_control_assessments %>% arrange(Participant)

# Add 'group' column with value 'training' for all rows
long_control_assessments$group <- 'control'

# Combine the two dataframes vertically
long_df <- bind_rows(long_active_assessments, long_control_assessments)


# Add Sex and Age data
merged_df <- merge(long_df, df[, c("Participant", "Age", "Sex")], by = "Participant", all.x = TRUE)
long_df <- merged_df
rm(merged_df)

shapiro.test(long_df$IP1)
shapiro.test(long_df$IP2)
shapiro.test(long_df$pS)

hist(long_df$IP1, main = "Histogram of IP1", xlab = "IP1 Values", ylab = "Frequency")
hist(long_df$IP2, main = "Histogram of IP2", xlab = "IP2 Values", ylab = "Frequency")
hist(long_df$pS, main = "Histogram of pS", xlab = "pS Values", ylab = "Frequency")


# Recode group: control = -1, training = 1
long_df$group <- ifelse(long_df$group == "control", -1, 1)

# Recode sex: male = -1, female = 1
long_df$Sex <- ifelse(long_df$Sex == "Male", -1, 1)

# Mean centre age
long_df[, "Age"] <- scale(long_df[, "Age"], center = TRUE, scale = FALSE)

long_df

```


```{r Change in IP1}

model <- lmer(formula = IP1 ~ group * time + Sex + Age + (1 | Participant), 
               data    = long_df)
summary(model)

# vif(model)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


emm1 = emmeans(model, specs = pairwise ~  time|group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="None")
#emm1$emmeans
#emm1$contrasts

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))

```


```{r Change in IP2}

model <- lmer(formula = IP2 ~ group * time + Sex + Age + (1 | Participant), 
               data    = long_df)
summary(model)

# vif(model)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")


emm1 = emmeans(model, specs = pairwise ~  time|group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="None")
#emm1$emmeans
#emm1$contrasts

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))

```


```{r Change in pS}

model <- lmer(formula = pS ~ group * time + Sex + Age + (1 | Participant), 
               data    = long_df)
summary(model)

# vif(model)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- EMMEANS -----------------------")

emm1 = emmeans(model, specs = pairwise ~  time|group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="None")
#emm1$emmeans
#emm1$contrasts

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))

```

# Explaining anxiety treatment response

## Computational parameters explain anxiety reduction in training group
```{r Prepare data for state LME}
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE) %>% 
  filter(Group == "training")

temp_df <- merge(temp_df, combined_HR, by = "Participant")

# No need to transform any predictors; residuals are normal
# temp_df <- applyOptLogTransform(temp_df, c(param_cols,"IP2_change", anx_cols))

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in state anxiety}

model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS +  State_BL + Sex + Age + mean_heartRate + mean_heartRateSD + (1 | Sex), 
               data    = temp_df,
              control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
isSingular(model, tol = 1e-4)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r Generate effects plot for state anxiety}

trellis.par.set(
  fontsize = list(
    text = 18,   # axis titles, strip labels
    points = 12  # tick mark labels
  )
)

est1 <- Effect("IP1", model, residuals=TRUE, KR=TRUE)
est2 <- Effect("pS", model, residuals=TRUE, KR=TRUE)


plt1 <- plot(est1,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,          # remove title
             partial.residuals = list(
               smooth = TRUE,
               pch = 16,
               col = "black",
               cex = 0.7,
               lty = "dashed"
             ),
             axes = list(
               x = list(IP1 = list(lab = bquote("Afferent signal precision (" ~ IP[1] ~ ")"))),
               y = list(lab = "Change in state anxiety", lim = c(-33, 33)),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)

plt2 <- plot(est2,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,          # remove title
             partial.residuals = list(
               smooth = TRUE,
               pch = 16,
               col = "black",
               cex = 0.7,
               lty = "dashed"
             ),
             axes = list(
               x = list(pS = list(lab = "Prior bias (pS)")),
               y = list(lab = "Change in state anxiety", lim = c(-33, 33)),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)



# Save figure
figures_path <- "figures_revision"

# Create folder if it doesn't exist
if (!dir.exists(figures_path)) {
  dir.create(figures_path, recursive = TRUE)
}
png(file.path(figures_path, "state_anx_effects_plot.png"), width = 600, height = 800)
grid.arrange(plt1, plt2, ncol = 1)
dev.off()


```


```{r Power analysis state anxiety}

# SESOI SIMULATION
SESOI <- fixef(model) * .75 # specify SESOI, 75% of effect sizes found

power_df <- temp_df %>%
  mutate(subject = as.numeric(Participant))

power_SESOI <- mixedpower(model = model, data = power_df,
                              fixed_effects = c("IP1", "pS"),
                              simvar = "subject", steps = c(28),
                              critical_value = 2, n_sim = 500,
                              SESOI = SESOI, databased = T, maxCores = 10)

power_SESOI

```

```{r Ridge regression state anxiety}

set.seed(67)

model <- linearRidge(temp_df$StateAnxietyChange_FmB ~ IP1 + IP2_change + pS + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD, data = temp_df)
summary(model)

```


```{r Prepare data for trait LME}

anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE) %>% 
  filter(Group == "training")

temp_df <- merge(temp_df, combined_HR, by = "Participant")

# Check predictors for normality and skew
shapiro.test(temp_df$IP1)
shapiro.test(temp_df$IP2_change)
shapiro.test(temp_df$IP2)
shapiro.test(temp_df$pS)
shapiro.test(temp_df$Age)
shapiro.test(temp_df$mean_heartRate)
shapiro.test(temp_df$mean_heartRateSD)


# For supplementary results: log-transforming IP2_change does not change the significant effects
# temp_df <- applyOptLogTransform(temp_df, c("IP2_change"))
# transformation_IP2_change <- optLogTransform(data.matrix(temp_df[, 'IP2_change']), type = "log", scaled = FALSE, skew_thresh = .5)

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in trait anxiety}

model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS + Trait_BL + Age + Sex + mean_heartRateSD + (1 | Sex), 
               data    = temp_df)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r Generate effects plot for trait anxiety}

trellis.par.set(
  fontsize = list(
    text = 18,  
    points = 12 
  )
)

# Create Effects
est1 <- Effect("IP1", model, residuals = TRUE, KR = TRUE)
est2 <- Effect("IP2_change", model, residuals = TRUE, KR = TRUE)
est3 <- Effect("pS", model, residuals = TRUE, KR = TRUE)

# Plot each effect
plt1 <- plot(est1,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,
             partial.residuals = list(
               smooth = TRUE, pch = 16, col = "black", cex = 0.7, lty = "dashed"
             ),
             axes = list(
               x = list(IP1 = list(lab = bquote("Afferent signal precision (" ~ IP[1] ~ ")"))),
               y = list(lab = "Change in trait anxiety", lim = c(-33, 33)),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)

plt2 <- plot(est2,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,
             partial.residuals = list(
               smooth = TRUE, pch = 16, col = "black", cex = 0.7, lty = "dashed"
             ),
             axes = list(
               x = list(IP2_change = list(lab = bquote("Change in precision weighting (" ~ Delta~IP[2] ~ ")"))),
               y = list(lab = "Change in trait anxiety", lim = c(-33, 33)),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)

plt3 <- plot(est3,
             lattice = TRUE,
             multiline = FALSE,
             main = NULL,
             partial.residuals = list(
               smooth = TRUE, pch = 16, col = "black", cex = 0.7, lty = "dashed"
             ),
             axes = list(
               x = list(pS = list(lab = "Prior bias (pS)")),
               y = list(lab = "Change in trait anxiety", lim = c(-33, 33)),
               grid = FALSE
             ),
             lines = list(col = c("black", "darkgrey"), lwd = 2)
)

# Save figure
figures_path <- "figures_revision"
if (!dir.exists(figures_path)) dir.create(figures_path, recursive = TRUE)

png(file.path(figures_path, "trait_anx_effects_plot.png"), width = 600, height = 1200)
grid.arrange(plt1, plt2, plt3, ncol = 1, heights = c(3, 3, 3))
dev.off()


```

```{r Power analysis trait anxiety}

SESOI <- fixef(model) * .75 # specify SESOI, 75% of effect sizes found

power_df <- temp_df %>%
  mutate(subject = as.numeric(Participant))

power_SESOI <- mixedpower(model = model, data = power_df,
                              fixed_effects = c("pS"), #"IP1", "IP2_change", "pS"
                              simvar = "subject", steps = c(28),
                              critical_value = 2, n_sim = 500,
                              SESOI = SESOI, databased = T, maxCores = 10)

power_SESOI

```

```{r Ridge regression trait anxiety}

set.seed(67)

model <- linearRidge(temp_df$TraitAnxietyChange_FmB ~ IP1 + IP2_change + pS + Trait_BL + Age + Sex + mean_heartRateSD, data = temp_df)
summary(model)

```

### In both groups pooled

```{r Prepare data for state LME}
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)

temp_df <- merge(temp_df, combined_HR, by = "Participant")

# No need to transform any predictors; residuals are normal
# temp_df <- applyOptLogTransform(temp_df, c(param_cols,"IP2_change", anx_cols))

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in state anxiety}

# LME doesn't converge
# model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS +  State_BL + Sex + Age + mean_heartRate + mean_heartRateSD + (1 |  Sex), 
#                data    = temp_df,
#               control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
# isSingular(model, tol = 1e-4)


model <- lm(formula = StateAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS +  State_BL + Sex + Age + mean_heartRate + mean_heartRateSD, 
               data    = temp_df,
              control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r Prepare data for trait LME}
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)
temp_df <- merge(temp_df, combined_HR, by = "Participant")

# Check predictors for normality and skew
shapiro.test(temp_df$IP1)
shapiro.test(temp_df$IP2_change)
shapiro.test(temp_df$IP2)
shapiro.test(temp_df$pS)
shapiro.test(temp_df$Age)
shapiro.test(temp_df$mean_heartRate)
shapiro.test(temp_df$mean_heartRateSD)


# For supplementary results: log-transforming IP2_change does not change the significant effects
# temp_df <- applyOptLogTransform(temp_df, c("IP2_change"))
# transformation_IP2_change <- optLogTransform(data.matrix(temp_df[, 'IP2_change']), type = "log", scaled = FALSE, skew_thresh = .5)

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in trait anxiety}

model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS + Trait_BL + Age + Sex + mean_heartRateSD + (1 | Sex), 
               data    = temp_df)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

### In control group only

```{r Prepare data for state LME}
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE) %>% 
  filter(Group == "control")

temp_df <- merge(temp_df, combined_HR, by = "Participant")

# No need to transform any predictors; residuals are normal
# temp_df <- applyOptLogTransform(temp_df, c(param_cols,"IP2_change", anx_cols))

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in state anxiety}

model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS +  State_BL + Sex + Age + mean_heartRate + mean_heartRateSD + (1 | Sex), 
               data    = temp_df,
              control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

isSingular(model, tol = 1e-4)
tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r Prepare data for trait LME}
anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)
temp_df <- merge(temp_df, combined_HR, by = "Participant")  %>% 
  filter(Group == "control")

# Check predictors for normality and skew
shapiro.test(temp_df$IP1)
shapiro.test(temp_df$IP2_change)
shapiro.test(temp_df$IP2)
shapiro.test(temp_df$pS)
shapiro.test(temp_df$Age)
shapiro.test(temp_df$mean_heartRate)
shapiro.test(temp_df$mean_heartRateSD)


# For supplementary results: log-transforming IP2_change does not change the significant effects
# temp_df <- applyOptLogTransform(temp_df, c("IP2_change"))
# transformation_IP2_change <- optLogTransform(data.matrix(temp_df[, 'IP2_change']), type = "log", scaled = FALSE, skew_thresh = .5)

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df %>% dplyr::select(Participant, IP1, IP2, IP2_change, pS, Group, Age, Sex, StateAnxietyChange_FmB, TraitAnxietyChange_FmB, mean_heartRate, mean_heartRateSD)
```


```{r LME to explain change in trait anxiety}

model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + IP1 + IP2_change + pS + Trait_BL + Age + Sex + mean_heartRateSD + (1 | Sex), 
               data    = temp_df)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

## Conventional task measures do not explain anxiety reduction (Supplementary 8.1 - 8.2)

See code at bottom of this cell to reproduce Supplementary 8.1.3 and 8.1.4 (using absolute values of d')
```{r Prepare data for LMEs}

anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE)

# Filter for training group only (Supplementary 8.1) or control group only (Supplementary 8.2)
# temp_df <- temp_df %>% filter(Group == "training")
temp_df <- temp_df %>% filter(Group == "control")


HR_change_df <- combined_HR_session %>%
  filter(Time %in% c("baseline", "final")) %>%
  pivot_wider(
    id_cols = Participant,
    names_from = Time,
    values_from = c(mean_heartRate, mean_heartRateSD),
    names_sep = "_"
  ) %>%
  mutate(
    mean_heartRate_change = mean_heartRate_final - mean_heartRate_baseline,
    mean_heartRateSD_change = mean_heartRateSD_final - mean_heartRateSD_baseline
  ) %>%
  dplyr::select(Participant, mean_heartRate_change, mean_heartRateSD_change)

temp_df <- merge(temp_df, combined_HR_assessment, by = "Participant")


# Calculate change in conventional task measures from baseline to final timepoints
temp_df$Dprime_change <- temp_df$Dprime_F - temp_df$Dprime_BL
temp_df$Cvalue_change <- temp_df$Cvalue_F - temp_df$Cvalue_BL
temp_df$HBT_change <- temp_df$HBT_FL - temp_df$HBT_BL

shapiro.test(temp_df$Dprime_change)
shapiro.test(temp_df$Cvalue_change)
shapiro.test(temp_df$HBT_change)

# Transform any non-normal variables
temp_df <- applyOptLogTransform(temp_df, c("Cvalue_change"))

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)


# For Reviewer 2: does using absolute values of d' allow for explanation of anxiety change outcomes?
# temp_df$Dprime_change <- abs(temp_df$Dprime_F) - abs(temp_df$Dprime_BL)

temp_df

```

```{r Multiple regression to predict change in state anxiety}

# For 8.1.1 and 8.2.1, LME has a non-singular Hessian, i.e., cannot trust beta estimates
model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + Dprime_change + Cvalue_change + HBT_change + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), data    = temp_df)
isSingular(model, tol = 1e-4)

# So for 8.1.1 and 8.2.1 we use this multiple regression model instead
# model <- lm(formula = StateAnxietyChange_FmB ~ 1 + Dprime_change + Cvalue_change + HBT_change + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD, data    = temp_df)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```


```{r LME to predict change in trait anxiety}
model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + Dprime_change + Cvalue_change + HBT_change + Trait_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), data    = temp_df)

isSingular(model, tol = 1e-4)


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```


## No prediction of treatment response using baseline measures

### Using baseline computational parameters

```{r Prepare data for LMEs}

temp_df <- training_B
temp_df <- merge(temp_df, df, by = "Participant", all.x = TRUE)
temp_df$Group <- ifelse(grepl("C", temp_df$Participant), "control", "training")

temp_df <- merge(temp_df, combined_HR_assessment, by = "Participant")

# Transform any non-normal variables from right dataframe
temp_df <- applyOptLogTransform(temp_df, c(param_cols,anx_cols,baseline_cols))

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)


# Recode group: control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

temp_df
```


```{r Trait anxiety change}
model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + IP1 + IP2 + pS + Trait_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), 
               data    = temp_df)
# summary(model)


summ(model)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

```

```{r State anxiety change}
model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + IP1 + IP2 + pS + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), 
               data    = temp_df)
# summary(model)


summ(model)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```

### Using baseline conventional interoceptive task parameters

```{r Prepare data for LMEs}

anx_cols <- c("TraitAnxietyChange_FmB", "StateAnxietyChange_FmB")
temp_df <- merge(combined_df, df, by = "Participant", all.x = TRUE) %>% 
  filter(Group == "training")

temp_df <- merge(temp_df, combined_HR_session %>% filter(Time == 'baseline'), by = "Participant")

shapiro.test(temp_df$Dprime_BL)
shapiro.test(temp_df$Cvalue_BL)
shapiro.test(temp_df$HBT_BL)

# Transform any non-normal variables
# temp_df <- applyOptLogTransform(temp_df, c("HBT_BL"))

# Recode group : control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)


temp_df

```

```{r State anxiety change}

# LME has singular Hession; coefficient estimates not trustworthy
# model <- lmer(formula = StateAnxietyChange_FmB ~ 1 + Dprime_BL + Cvalue_BL + HBT_BL + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), data    = temp_df)

model <- lm(formula = StateAnxietyChange_FmB ~ 1 + Dprime_BL + Cvalue_BL + HBT_BL + State_BL + Age + Sex + mean_heartRate + mean_heartRateSD, data    = temp_df)

summ(model)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
# qqnorm(residuals)
# qqline(residuals)
```


```{r Trait anxiety change}

model <- lmer(formula = TraitAnxietyChange_FmB ~ 1 + Dprime_BL + Cvalue_BL + HBT_BL + Trait_BL + Age + Sex + mean_heartRate + mean_heartRateSD + (1 | Sex), data    = temp_df)
isSingular(model, tol = 1e-4)

summ(model)

tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- EMTRENDS -----------------------")

# 
# emt <- emtrends(model,specs=pairwise~Sex,var="IP2")
# summary(emt,infer=TRUE)

print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk statistic =", round(shapiro_test$statistic, 2), ", p-value =", round(shapiro_test$p.value, 2), "\n")

# Check VIFs for multicollinearity
vifs <- vif(model)
cat("Variance inflation factors:\n")
vifs <- round(vifs, 2)
cat(paste(names(vifs), vifs, sep = ": ", collapse = ";  "))


# Q-Q plot of the residuals
# qqnorm(residuals)
# qqline(residuals)
```

# MAIA Self-reported interoception (exploratory analyses)

```{r Prepare data for LMEs}

MAIA_change_cols = c("Change_MAIA_Noticing_FLmBL", "Change_MAIA_NotDistracting_FLmBL","Change_MAIA_NotWorrying_FLmBL", "Change_MAIA_AttentionReg_FLmBL", "Change_MAIA_EmoAware_FLmBL", "Change_MAIA_SelfReg_FLmBL", "Change_MAIA_BodyList_FLmBL", "Change_MAIA_Trusting_FLmBL")

MAIA_BL_cols <- c('NoticingBL', 'not_distractingBL', 'not_worryingBL', 'attention_regulationBL','emotional_awarenessBL', 'self_regulationBL', 'body_listeningBL', 'trustingBL')
MAIA_FL_cols <- c('NoticingFL','not_distractingFL','not_worryingFL','attention_regulationFL','emotional_awarenessFL','self_regulationFL','body_listeningFL','trustingFL')

temp_df <- df[,c('Participant', 'Age', 'Sex', MAIA_change_cols, MAIA_BL_cols, MAIA_FL_cols, "StateAnxietyChange_FmB","TraitAnxietyChange_FmB")]

temp_df <- merge(combined_df, temp_df, by = "Participant", all.x = TRUE)

# Recode group: control = -1, training = 1
temp_df$Group <- ifelse(temp_df$Group == "control", -1, 1)

# Recode sex into sum coding: male = -1, female = 1
temp_df$Sex <- ifelse(temp_df$Sex == "Male", -1, 1)

# Mean centre age
temp_df[, "Age"] <- scale(temp_df[, "Age"], center = TRUE, scale = FALSE)

# Calculate total MAIA baseline score
temp_df$MAIA_totalBL <- rowSums(temp_df[, MAIA_BL_cols], na.rm = TRUE)

# Calculate total MAIA final score
temp_df$MAIA_totalFL <- rowSums(temp_df[, MAIA_FL_cols], na.rm = TRUE)

# Tidy up column names
names(temp_df) <- gsub("_FLmBL$", "", names(temp_df))

temp_df

```

This block runs an LME for one subscale; change the variable 'col' to change the subscale to analyse.
```{r Change in MAIA subscales}

col <- 6 # Change from 1 through 8 to run LME on each subscale

col_name <- str_sub(MAIA_BL_cols[col],1,-3)
# Convert to long form
long_df <- temp_df[,c('Group', 'Participant', 'Age', 'Sex', MAIA_BL_cols[col], MAIA_FL_cols[col])]
long_df <- pivot_longer(long_df, cols = starts_with(col_name), names_to = "Time", values_to = col_name)

long_df <- long_df %>%
  mutate(Time = ifelse(grepl("BL$", Time), "baseline", 
                                ifelse(grepl("FL$", Time), "final", Time)))
# Merge HR and HRV data
long_df <- long_df %>%
  left_join(combined_HR_session, by = c("Participant", "Time"))

long_df



model <- lmer(paste(col_name, "~ Group + Time + Age + Sex + Group:Time +  (1 | Participant)"), data = long_df) 


tab_model(model, show.ci = FALSE, show.se = TRUE, show.df = TRUE, p.val = "kr", show.stat = TRUE)

print("----------------------- EMMEANS -----------------------")


emm1 = emmeans(model, specs = pairwise ~  Time|Group, lmer.df = "kenward-roger", calc = c(n = ".wgt."), weights = "prop")
summary(emm1, adjust="holm")
#emm1$emmeans
#emm1$contrasts


print("----------------------- Normality of residuals -----------------------")

# Residuals from the model
residuals <- residuals(model)

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals)

cat("Shapiro-Wilk Test:\n")
cat("Statistic: ", shapiro_test$statistic, "\n")
cat("p-value: ", shapiro_test$p.value, "\n")

# Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)

print("----------------------- Effect size -----------------------")
eff_size(emm1, sigma = sigma(model), edf = df.residual(model))
```

```{r Exploratory correlations with anxiety reduction}

temp_df <- temp_df %>% 
  dplyr::filter(Group == 1) %>% 
  dplyr::select(Participant, Group, starts_with("change_MAIA"),StateAnxietyChange_FmB,TraitAnxietyChange_FmB, IP1, IP2_change, pS)

MAIA_change_cols <- temp_df %>% 
  dplyr::select(starts_with("change_MAIA")) %>% 
  colnames()

# Computational parameters
cor.test(temp_df$pS, temp_df$Change_MAIA_Noticing, method = "pearson")
cor.test(temp_df$pS, temp_df$Change_MAIA_Trusting, method = "pearson")

# Anxiety change
cor.test(temp_df$TraitAnxietyChange_FmB, temp_df$Change_MAIA_AttentionReg, method = "pearson")
cor.test(temp_df$StateAnxietyChange_FmB, temp_df$Change_MAIA_NotWorrying, method = "pearson")


```

```{r Correlation scatterplots (Figure 11)}

# 2 horizontal subplots
# Use consistent black and white formatting

# Change in MAIA Noticing (y) and pS (x), plt1
plt1 <- ggplot(temp_df, aes(x = pS, y = Change_MAIA_Noticing)) +
  geom_point(size = 3, color = "black") +
  geom_smooth(method = "lm", color = "black", se = TRUE, fill = "lightgrey", size = 1) +
  theme_minimal() +
  theme(
    text = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black")
  ) +
  labs(
    x = "Prior bias (pS)",
    y = "Change in MAIA Noticing"
  )

# Change in MAIA Trusting (y) and pS (x), plt2
plt2 <- ggplot(temp_df, aes(x = pS, y = Change_MAIA_Trusting)) +
  geom_point(size = 3, color = "black") +
  geom_smooth(method = "lm", color = "black", se = TRUE, fill = "lightgrey", size = 1) +
  theme_minimal() +
  theme(
    text = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black")
  ) +
  labs(
    x = "Prior bias (pS)",
    y = "Change in MAIA Trusting"
  )



# Save figure
figures_path <- "figures_revision"

# Create folder if it doesn't exist
if (!dir.exists(figures_path)) {
  dir.create(figures_path, recursive = TRUE)
}
png(file.path(figures_path, "Figure_11.png"), width = 800, height = 400)
grid.arrange(plt1, plt2, ncol = 2)
dev.off()

```
